### 2019-02-28 papers/journal/encoding.tex
Updated 6.3

In order to specify the stability condition, we use a combination of enumeration and nondeterminism.
%
Nondeterminism can be viewed as an oracle that encodes a correct $m$-path for a specific $m$ if it exists. 
%
We then use this path as a witness to ensures that stability is maintained for each outgoing molecule $m$ on every edge $E$.
%

\subsubsection{Constraints for stability condition}
%
We use \texttt{NOracle} to represent the required path to encode the stability condition in VTS.
%
A naive implementation is shown in~\ref{eq:reachbmc1} where \texttt{NOracle}$(m, i, j, w)$ represent the presence of a $m$-$path$ from node $i$ to node $j$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
It performs enumeration of every $m$-$path$, $m$-${paths}_i^j$, of at most $w$ length. 
%
Exploring $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on it, until $m$-${paths}_i^j$ of length $w-1$.
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.

@@

We encode stability using \texttt{NOracle} and if there is an $m$-edge between node $i$ and $j$, there has to be a $m$-path from $j$ to $i$ of length maximum w.
%
We have used bv $\unrhd$ m to extract the $m^{th}$ bit of the bitvector bv.

\begin{align}
 \bigwedge\limits_{i,j,m} (\bigvee_{q} (\texttt{bvEdge}[i][j][q] \unrhd m) := 1) \limplies NOracle(m, j, i, w)
  \tag{S2}\label{eq:reachbmc2}
\end{align}

@@

\subsection{The key improvement}
The key improvement of SMT tool over the BMC is the
encoding of reachability.
% which was done using an enumeration of paths. 
%
In SMT tool, reachability is encoded in two different
ways in constraints~\eqref{eq:reach1} and~\eqref{eq:drop3}.
%
The reachability is recursively defined in~\eqref{eq:drop3} and has
trivial solutions by making all $r'$s true.
%
However, the trivial solutions are disallowed by constraint~\eqref{eq:drop4} and we find
only the solutions that captures the evidence of unreachability.
%
On the other hand,
we have added length of paths in our reachability encoding in constraint~\eqref{eq:reach1},
which needs relatively more auxiliary variables.
%
This is because constraint~\eqref{eq:reach2} has only positive
occurrences of the reachability variables and if we had defined
$r$s in ~\eqref{eq:reach1} without paths,
the circular dependencies in the recursive definitions of $r$s
may have resulted in spurious satisfying assignments that
do not encode reachability.
%
By adding the path length, we break the circular dependencies, the
constraint remains polynomial in size, and satisfying assignments only
correspond to correct reachability.

>>>>

In order to specify the stability condition, we use a combination of enumeration and nondeterminism.
%
Nondeterminism can be viewed as an oracle that guesses a correct $m$-path for a specific $m$ if it exists. 
%
We then use this path as a witness to ensures that stability is maintained for each outgoing molecule $m$ on every edge $E$.
%
\subsubsection{Constraints for stability condition}
%
We use \texttt{NOracle} to guess the required path to encode the stability condition in VTS.
%
A naive implementation is shown in~\ref{eq:reachbmc1} where \texttt{NOracle}$(m, i, j, w)$ represent the presence of a $m$-$path$ from node $i$ to node $j$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
It performs enumeration of every $m$-$path$, $m$-${paths}_i^j$, of at most $w$ length. 
%
Starting with exploring $m$-${paths}_i^j$ of length 1, a direct edge between $i$ and $j$ with molecule $m$ present on it, until every $m$-${paths}_i^j$ of length $w-1$ is explored.
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.

@@

We encode stability using \texttt{NOracle}. If there is an $m$-edge between node $i$ and $j$, there has to be a $m$-$path_i^j$ from $j$ to $i$ of maximum $w$ length.
%
The expression (\texttt{bv} $\unrhd$ $m$) extracts the $m^{th}$ bit of the bitvector \texttt{bv}.

\begin{align}
	\bigwedge\limits_{i,j,m} (\bigvee_{q} (\texttt{bvEdge}[i][j][q] \unrhd m) := 1) \limplies \texttt{NOracle}(m, j, i, w)
  \tag{S2}\label{eq:reachbmc2}
\end{align}

@@

\subsection{The key difference in the reachability encoding}
The key improvement of the SMT tool over the BMC is the
encoding of reachability.
% which was done using an enumeration of paths.
%
In the BMC tool, reachability is encoded by enumerating all the possible paths using nondeterminism.
%
This encoding is non-optimal in the sense that in the worst case an exponential number of paths needs to be explored.
%
Whereas in the SMT tool, reachability is encoded in two different
ways in constraints~\eqref{eq:reach1} and~\eqref{eq:drop3}.
%
The reachability is recursively defined in~\eqref{eq:drop3} and has
trivial solutions by making all $r'$s true.
%
However, the trivial solutions are disallowed by constraint~\eqref{eq:drop4} and we find
only the solutions that captures the evidence of unreachability.
%
On the other hand,
we have added length of paths in our reachability encoding in constraint~\eqref{eq:reach1},
which needs relatively more auxiliary variables.
%
This is because constraint~\eqref{eq:reach2} has only positive
occurrences of the reachability variables and if we had defined
$r$s in ~\eqref{eq:reach1} without paths,
the circular dependencies in the recursive definitions of $r$s
may have resulted in spurious satisfying assignments that
do not encode reachability.
%
By adding the path length, we break the circular dependencies, the
constraint remains polynomial in size, and satisfying assignments only
correspond to correct reachability.
